---
title: "Session 2: Homework 1"
author: "Group 8"
date: "`r Sys.Date()`"
output:
  html_document:
    theme: flatly
    highlight: zenburn
    number_sections: yes
    toc: yes
    toc_float: yes
    code_folding: show
---

```{r, setup, echo=FALSE}
knitr::opts_chunk$set(
  message = FALSE, 
  warning = FALSE, 
  tidy=FALSE,     # display code as typed
  size="small")   # slightly smaller font for code
options(digits = 3)

knitr::opts_chunk$set(
  fig.width=8, 
  fig.height=6,
  fig.align = "center"
)
```

```{r load-libraries, warning=FALSE, message=FALSE, echo=FALSE}
library(tidyverse)  # Load ggplot2, dplyr, and all the other tidyverse packages
library(mosaic)
library(ggthemes)
library(lubridate)
library(here)
library(skimr)
library(janitor)
library(vroom)
library(tidyquant)
library(dplyr)
library(ggplot2)
```
# Analysis of movies- IMDB dataset

```{r,load_movies, warning=FALSE, message=FALSE}
# Assign IMDB data to variable
movies <- read_csv(here::here("data", "movies.csv"))

```

## Use your data import, inspection, and cleaning skills to answer the following:

-   Are there any missing values (NAs)? Are all entries distinct or are there duplicate entries?

There are no missing values according to the skim function. Nevertheless, we notice some duplicate values, for example "Nightmare on Elm Street" or "Alice in Wonderland". We have to be careful to look at more than just the title in identifying duplicates, as some movies could have the same name, but be released twice on different dates or have a different director.

```{r,Skim_data}
# Skim IMDB data
skim(movies)

```

-   Produce a table with the count of movies by genre, ranked in descending order

    ```{r,count_of_movies}
    movies%>%
      count(sort = TRUE,genre)
    ```

-   Produce a table with the average gross earning and budget (`gross` and `budget`) by genre. Calculate a variable `return_on_budget` which shows how many \$ did a movie make at the box office for each \$ of its budget. Ranked genres by this `return_on_budget` in descending order

    ```{r,avg_gross_budget}
    # Create table with the average gross earning and budget
    avg_genre = movies%>%
      group_by(genre) %>% 
      summarize(avg_earning = sum(gross)/count(genre),
                avg_budget = sum(budget)/count(genre)) # Create 2 columns to store the average earning and budget
    avg_genre%>% 
      mutate(return_on_budget = avg_earning/avg_budget) %>% # The return is just the earning/budget
      arrange(desc(return_on_budget))
    ```

-   Produce a table that shows the top 15 directors who have created the highest gross revenue in the box office. Don't just show the total gross amount, but also the mean, median, and standard deviation per director.

    ```{r,top_15}
    # Calculate summary statistics for top 15 directors 
    Top_directors = movies%>%
      group_by(director) %>% 
      summarise(sum_gross = sum(gross),
                mean_gross = mean(gross),
                median_gross = median(gross),
                SD_gross = sd(gross))
    # Choose the top 15 with highest gross earnings
    Top_directors%>%
      slice_max(sum_gross,n = 15)
    ```

-   Finally, ratings. Produce a table that describes how ratings are distributed by genre. We don't want just the mean, but also, min, max, median, SD and some kind of a histogram or density graph that visually shows how ratings are distributed.

    ```{r,rating}
    # Calculate summary statistics for ratings by genre
    Ratings_genre = movies%>%
      group_by(genre)%>%
      summarise(mean_ratings = mean(rating),
                min_rating = min(rating),
                max_rating = max(rating),
                median_rating = median(rating),
                SD_rating = sd(rating))
    Ratings_genre

    # Plot ratings by genre
    ggplot(movies,aes(x=rating)) +
      geom_density() +
      labs(title="There are few completely unpopular movies with a rating of less than 5",subtitle = "Density plot of movie ratings on IMDB",x = "Rating", y = "Density")+
      theme_bw()
    ```

## Use `ggplot` to answer the following

-   Examine the relationship between `gross` and `cast_facebook_likes`. Produce a scatterplot and write one sentence discussing whether the number of facebook likes that the cast has received is likely to be a good predictor of how much money a movie will make at the box office. What variable are you going to map to the Y- and X- axes?

While there seems to be a minor correlation between the amount of cast facebook likes and the money a movie makes, the relationship is not strong enough to make it a good predictor of a movie's success.

```{r, gross_on_fblikes}
# Map revenue vs Facebook likes
movies%>%
ggplot(aes(x=cast_facebook_likes, y = gross)) +
  geom_point()+scale_x_log10()+geom_smooth(method = "lm", se = FALSE) +
  labs(title="Cast Facebook likes do not seem to be a reliable predictor of movie success",subtitle = "Scatterplot of number of cast facebook likes and movie gross revenue",x = "Number of Cast Facebook Likes", y = "Gross Revenue")+
  theme_bw()
```

-   Examine the relationship between `gross` and `budget`. Produce a scatterplot and write one sentence discussing whether budget is likely to be a good predictor of how much money a movie will make at the box office.

While a budget is surely not a guarantor of movie success , the fitted line has a positive slope and therefore implies that with rising budget comes rising gross revenue.

```{r, gross_on_budget}
 #Map revenue vs budget
movies%>%
  ggplot(aes(x=budget,y=gross))+
  geom_point()+geom_smooth(method = "lm", se = FALSE) +
  labs(title="A higher budget seems to positively affect movie gross revenue",subtitle = "Scatterplot of movie budget and movie gross revenue",x = "Budget", y = "Gross Revenue")+
  theme_bw()

```

-   Examine the relationship between `gross` and `rating`. Produce a scatterplot, faceted by `genre` and discuss whether IMDB ratings are likely to be a good predictor of how much money a movie will make at the box office. Is there anything strange in this dataset?

Generally, higher ratings indicates higher gross earnings for all genres for which we have a significant amount of data. However, we can also see that it is possible for movies to have a good rating while not making a lot of money. This most likely concern the likes of indie movies, that receive strong support but never make it into the pop culture.

There are some interesting anomalies in the data in the form of extreme values. Drama movies usually fall within the same range of gross revenues, however "Titanic" by James Cameron reports much higher values than the rest. Coincidentally, James Cameron is also the director for highest grossing movie in the data "Avatar". It is also noticeable that there are no observable outliers for genres like comedy or adventure, even though there are several movies of the genre in the data. Apparently these categories do not display the required parameters to polarize the nation.

```{r, gross_on_rating}
# Map revenue vs rating, faceted by genre
movies%>%
  ggplot(aes(x=rating, y = gross,color=genre))+
  geom_point()+facet_wrap(~genre)+
  labs(title="A higher rating seems to be correlated to higher gross revenues",subtitle = "Faceted scatterplot of IMDB rating and movie gross revenue",x = "IMDB Rating", y = "Gross Revenue")
```